Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
improvements to figures

We apply \system to three domains:
list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then constructs
its own DSL for expressing solutions in the domain (Fig.~\ref{initialExampleDSL}).
\begin{figure}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      % \code{[99\, 4\, 2]}$\to $\code{99}\\
      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[1]}$\to $\code{[1\, 3]}\\
      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

      \\\bottomrule

\end{tabular}}
\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them. Bottom: Subset of learned DSL. Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
\end{figure}

@@

We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Fig.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

\begin{figure}[b]
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (a b) (foldr b (cons "," a)}\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}\\
      \code{($\lambda$ (a b) (foldr a b }\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(foldr a b ($\lambda$ (x z) }\\
      \phantom{\code{(foldr a b }}\code{(cons x z)))}
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes fragments of programs to add to the DSL.
    These fragments are taken from programs in the frontiers (left column). Here, the proposed subexpression (top) appends two lists or (bottom) performs character substitutions.} 
  \label{fragmentExample}
\end{figure}

@@

\begin{wrapfigure}{r}{6.4cm} \newcommand{\functionSize}{1.5cm}\vspace{-0.5cm}\centering
%  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
  %% \includegraphics[width = \functionSize]{figures/functions/432.png}
  %% \includegraphics[width = \functionSize]{figures/functions/415.png}
  %% \includegraphics[width = \functionSize]{figures/functions/426.png}
  %% \includegraphics[width = \functionSize]{figures/functions/597.png}
  %% \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{l}{0.35\textwidth}\centering
\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
  \hspace{0.5cm}\emph{($f_2$: quadratics)}\\
  $f_3($\code{x}$)\,=\,$\code{(/ (}$f_2$\code{ x) (}$f_0$\code{ x))}\\
    \hspace{0.5cm}\emph{($f_3$: ratio of polynomials)}
  \\\bottomrule
\end{tabular}
\caption{Some learned subroutines for symbolic regression. System starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
\end{wrapfigure}

@@

@@

\begin{wrapfigure}{r}{0.4\textwidth}
  \begin{center}
    \includegraphics[width = 0.4\textwidth]{figures/geomCompiled.eps} 
  \end{center}
  \caption{
  Top left: hand crafted traning targets. Top right: examples of
  discovered compiled new programs. Bottom: compiled programm across iterations
  to highlight structure emergence.
  }\label{geomCompiled} 
\end{wrapfigure}

We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
domain-specific primitives that the algorithm itself discovers,
together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration
of top-down symbolic representations and bottom-up neural networks --
both of them learned -- could help make program induction systems more
generally useful for AI. Many directions remain open. Two immediate
goals 
are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in
some domains over the generic methods used here.
Another direction is to
explore DSL meta-learning: Can we find a \emph{single} universal
primitive set that could effectively bootstrap DSLs for new domains,
including the three domains considered,  but also many others?

>>>>

We apply \system to three domains:
list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then constructs
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{table}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      % \code{[99\, 4\, 2]}$\to $\code{99}\\
      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[1]}$\to $\code{[1\, 3]}\\
      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

  \end{tabular}
  \\\bottomrule 
\end{tabular}}\\\\
\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them. Bottom: Subset of learned DSL. Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
\end{table}

@@

We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Tbl.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

\begin{table}[b]
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (a b) (foldr b (cons "," a)}\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}\\
      \code{($\lambda$ (a b) (foldr a b }\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(foldr a b ($\lambda$ (x z) }\\
      \phantom{\code{(foldr a b }}\code{(cons x z)))}
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x '-') ',' x))) s)}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes fragments of programs to add to the DSL.
    These fragments are taken from programs in the frontiers (left column). Here, the proposed subexpression (top) appends two lists or (bottom) performs character substitutions.} 
  \label{fragmentExample}
\end{table}

@@

%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Tbl.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.

@@

\begin{figure*}[h]
\makebox[\textwidth][c]{
\begin{minipage}{0.55\textwidth} \newcommand{\functionSize}{1.5cm}\centering
%  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
  %% \includegraphics[width = \functionSize]{figures/functions/432.png}
  %% \includegraphics[width = \functionSize]{figures/functions/415.png}
  %% \includegraphics[width = \functionSize]{figures/functions/426.png}
  %% \includegraphics[width = \functionSize]{figures/functions/597.png}
  %% \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{minipage}
\hspace{1em}
\begin{minipage}{0.5\textwidth}
  \centering\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
  \hspace{0.5cm}\emph{($f_2$: quadratics)}\\
  $f_3($\code{x}$)\,=\,$\code{(/ (}$f_2$\code{ x) (}$f_0$\code{ x))}\\
    \hspace{0.5cm}\emph{($f_3$: ratio of polynomials)}
  \\\bottomrule
\end{tabular}
  \captionof{table}{Some learned subroutines for symbolic regression. System starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
\end{minipage}
}
\end{figure*}

@@

\parbox{0.58\textwidth}{
We contribute an algorithm, \systemEnding, that learns to program by bootstrapping a DSL with new
domain-specific primitives that the algorithm itself discovers,
together with a neural recognition model that learns how to
efficiently deploy the DSL on new tasks. We believe this integration
of top-down symbolic representations and bottom-up neural networks --
both of them learned -- could help make program induction systems more
generally useful for AI. Many directions remain open. Two immediate
goals 
are to integrate more sophisticated neural recognition
models~\cite{devlin2017robustfill} and program
synthesizers~\cite{solar2008program}, which may improve performance in
some domains over the generic methods used here.
Another direction is to
explore DSL meta-learning: Can we find a \emph{single} universal
primitive set that could effectively bootstrap DSLs for new domains,
including the three domains considered,  but also many others?
}\hfill\begin{minipage}{0.4\textwidth}\centering
  \includegraphics[width=0.8\textwidth]{figures/geomCompiled.eps} 
  \captionof{figure}{
  Top left: hand crafted traning targets. Top right: examples of
  discovered compiled new programs. Bottom: compiled programm across iterations
  to highlight structure emergence.
  }\label{geomCompiled}
\end{minipage}
